Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset

### 2025-06-27 code/manuscript.Rmd
updated supplementary material and manuscript

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events (step changes or more gradual changes in how a building operates) occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a meidum of 74% accuracy improvement and is even faster. Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability. We also assessed the impact of normalizing for different weather, starting the M&V at different time of the year, continuing randomization with a different sampling ratio after satisfying all criteria, dropping samples affected by carryover when switching between strategies, and identified the optimal sampling interval for that scenario using this large dataset.

@@

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level, where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. Typically, the process begins with
baseline measurements taken during the year before implementing the
energy-efficiency retrofit, followed by the same measurement procedure
during the year after the intervention. After collecting two years of
data, an M&V analyst fits an energy prediction model, using variables
such as outdoor temperature and time [@mathieu_quantifying_2011] to
project baseline energy consumption in the post-retrofit period. The
difference between the counter-factual baseline and the measured
intervention period energy consumption represents the energy savings.
One of the key drawbacks of this method is its reliance on a two-year
timeline to quantify savings, during which baseline measurements can
become outdated due to changes in building performance caused by
non-routine events unrelated to the intervention. Specifically, an M&V
analyst should account for two types of baseline changes: (1) static
changes, which are known building operational changes: such as
renovation, equipment addition or removal; and (2) gradual changes,
which often occur subtly and unknown, such as the incremental adoption
of LED lighting or more efficient plug loads. Identifying and
quantifying these changes are particularly labor-intensive and
challenging. Consequently, those limitations reduce the feasibility of
M&V and complicate the quantification of estimation uncertainty, thus
impacting the accuracy and timeliness of savings assessments.

@@

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston2019; @duan2013; @gupta2019;
@mie2022; @munro; @raseduzzaman2017]. An early application of this
randomized design to the field of M&V in buildings was first used in
[@raftery2018], which compared the energy performance of two different
supply air temperature reset strategies in a building. Subsequent
improvements to the M&V method include adding a sequential evaluation
framework and defining stopping criteria to end the M&V period early
after reaching a target level of uncertainty, among other improvements,
such as using blocked randomization by weekday. The full framework is
detailed in a previously published study [@raftery2024] along with a
case study application to a real building, and other example use cases.
In summary, this method provides M&V analysts with a randomized schedule
that alternates between baseline and intervention implementation while
ensuring balanced sampling across days of the week and seasons. For
example, given a 10-week M&V period for one intervention, the balanced
randomized schedule would equally sample five Mondays with the baseline
strategy in operation and five Mondays with the intervention strategy in
operation. The major limitation of this method is that it is only
applicable to a subset of retrofit projects where interventions can be
easily switched on and off, which is a small subset of all the
interventions performed in buildings. However, for all applicable use
cases, it allows analysts to detect energy savings sequentially shortly
after the test begins. In addition, once the desired estimate
uncertainty target is achieved, analysts can terminate the M&V study and
switch to 100% intervention. The key advantage of randomization is that
if control strategies are sampled with equal probability, the influence
of other confounding factors such as occupancy change (a common
non-routine event) and other more subtle long term changes in building
energy use are likely to be evenly distributed among measurements,
leading to a more accurate and unbiased assessment of the intervention
effect.

@@

Overall, those non-routine events mostly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. If M&V analysts are unaware of such changes and
lack an appropriate adjustment method (e.g., replacing filters before
the intervention begins), the savings could be overestimated as the
decreased energy use is incorrectly attributed to the intervention
rather than the filter replacement. An operational drift has not been
formally defined in the literature or standards, but in this study, we
consider it as a special type of non-routine event, as it also
represents a change in facility operation that occurs gradually over a
longer period of time.

@@

For all qualified buildings, we also create a feature to identify
buildings with very stable electricity usage overall by identifying
those with no statistically significant difference (P-value \> 0.05)
between the two-year electricity usage. Applying this filter to real
buildings, only 66 buildings are labeled in this 'stable' subset out of
a total of more than 600 buildings. While we did not adjust for weather
differences between the two years when generating this high-level
feature, this result highlights that the assumption that buildings are
typically stable over the timeframe involved in conventional M&V methods
is rarely true. As mentioned in \@ref(non-routine-events-impact), such
variability is largely associated with non-routine events, especially
operational drifts since the assessment, even under ideal situations,
lasts for two years. And according to the statistics shown here, it is a
typical case observed in real buildings applications.

@@

| Parameter              | Description                                                      | Value |
|------------------|------------------------------------|------------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

-   A minimum and maximum of 12 and 108 weeks respectively. The
    randomized schedule covers the entire two-year period but the
    stopping criteria enable an early stop at the end of the satisfied
    blocking period.

@@

The accuracy ($Acc$) is calculated as: $$
Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage and $E'$ indicates the
estimated electricity usage either through the conventional method or
the new randomized method. We present the distribution of savings
estimation accuracy using boxen plots, also known as letter plots, which
is an advanced variation of the box plot designed to extend beyond the
interquartile range (IQR) by progressively dividing the data into
smaller percentiles, revealing more detail in the tails of the
distribution. We set the division parameter to k = 4, meaning the entire
box area represents the data distribution from 6.25% to 93.75%. As the
steps move closer to the center line (50% median), the distribution
range progressively narrows with the next step representing 12.5% to
87.5%, followed by 25% to 75%. In addition, for each boxen plot, we also
showed the mean value of the distribution on top for reference.

@@

Additionally in the literature, we found another M&V study using over
500+ commercial buildings (on a different open-source dataset), but
their study scope was to compare the predictive performance of a variety
of M&V baseline models from simple weekly mean to more complex machine
learning methods using the normalized mean bias error (NMBE)
[@granderson_accuracy_2016]. Since we also calculated the normalized
savings estimation error, we compared our normalized accuracy for 600+
buildings with the TOWT model and TMY weather with their TOWT model
prediction assessment in the following table, highlighting that the
model and overall building datasets are performing similarly.

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-------------------|-----------------------------------|-------------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

@@

In some cases, particularly those involving large thermal mass and
active thermal energy storage systems or operating continuously, there
can be a substantial time-lagged effect associated with interventions.
This means that the effect of one control strategy may persist, or
'carryover', after switching to another. To mitigate such carryover
effects, building analysts may choose to drop non-consecutive days in
the switchback experiment. For example, consider a heat pump water
heater where the tank is charged at very different times of the day by
the intervention strategy compared to the baseline strategy. When
switching strategy, the tank may be much warmer or cooler than typical
for that strategy, thus influencing subsequent measurements,
particularly the performance measured on the first day (i.e., a
non-consecutive day). In such cases, increasing the sampling interval
(e.g., sample strategies every 2 or 3 days instead of daily) and
discarding measurements from those non-consecutive days can eliminate
carryover. However, this approach reduces the total data collected: for
instance, sampling every 3 days and dropping days with non-consecutive
control strategies would drop 1/6 of measurements on average. Sampling
every 2 days means dropping a quarter of the measurements. Therefore, by
increasing the sampling interval, there is an accuracy penalty due to a
reduction in randomization (how often samples occur), and when
decreasing the sampling interval, there is an accuracy penalty from
dropping more non-consecutive days. To quantify such trade-off, Figure
\@ref(fig:absinterval) shows the accuracy of different sampling
intervals and the consequence of dropping the non-consecutive days in
the dataset at the time when randomized M&V produces a result (after
satisfying all stopping criteria).

@@

```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=6, fig.width=12}

@@

In this study, we set the baseline period to begin on January 1st, which
offers the advantage of capturing colder outdoor weather conditions
early in the M&V process. However, when the baseline measurements miss
the initial winter period, meeting the required weather condition
criteria can potentially take a longer time. To assess this impact, we
repeated the randomized M&V analysis for all buildings using a new start
date in March. As shown in Figure \@ref(fig:startT), the new M&V process
results in a noticeable delay in satisfying the stopping criteria.
Compared to Figure \@ref(fig:intervaltime), where a January start
allowed most buildings to finish within 36 weeks, a March start extended
the required duration to 48 weeks for the majority of buildings.
Furthermore, when a longer sampling interval is used or non-consecutive
days are excluded (to avoid carryover effects), some buildings may
require up to 60 weeks to complete the randomized M&V. As expected, to
collect sufficient data representing winter conditions, the sampling may
need to continue until the next January (around week 48). However, as
the overall data collection period extends, the influence of dropping
non-consecutive days diminishes and becomes less of a concern.

@@

Specifically, we showed that throughout 11 different locations assessed
in this study, the randomized M&V can provide a saving estimation after
36 weeks (with the majority finished by 24 weeks) once all stopping
criteria are satisfied. In contrast, the conventional method requires a
full range of baseline and intervention measurements under normal
operating conditions, typically taking 6 - 9 months for each phase. Most
importantly, we verified that with a reduced M&V timeline, the
randomized method can estimate savings more accurately by showing that
the absolute error is only 1 - 2% for a typical building. Whereas for
the much longer two-year conventional method, the estimation error was
approximately 5% in a typical building. We also evaluated the impact of
non-routine events, especially more gradual and subtle operational drift
on the proposed M&V method and the results show that baseline changes in
the post-retrofit period can deviate savings estimated using the
conventional method. In contrast, we found that those events have a very
negligible impact on the savings estimated using the randomized method,
demonstrating robustness to this issue. Furthermore, we consider
scenarios where there is a known carryover effect from switching between
strategies in the building and assess the impact of dropping samples
where the strategy is non-consecutive and varying the sampling interval.
Through that process, we showed that for the typical building, if
carryover is present, using a 2-day sampling interval and dropping days
with non-consecutive control strategies operating yields the optimal
design for the median building in this dataset. Lastly, we investigated
the time required to complete the randomized M&V method under different
scenarios. We found that measuring a sufficient range of outdoor weather
conditions remains the most stringent criterion. But even for the most
non-ideal case, the randomized method can still produce an estimate much
faster than 2 years, and produces an estimate with far lower error.

>>>>

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^, Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events (step changes or more gradual changes in how a building operates) occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded. In this study, we demonstrated that switchable interventions, such as most control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 2 years). Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability. We also assessed the impact of normalizing for different weather, starting the M&V at different dates of the year, continuing randomization with a different sampling ratio after satisfying all criteria, and dropping samples affected by carryover effects when switching between strategies. For each scenario, we identified the optimal sampling interval using the large dataset.

@@

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level, where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. In conventional practice, the process
begins with baseline measurements taken during the year before
implementing the energy-efficiency retrofit, followed by the same
measurement procedure during the year after the intervention. After
collecting two years of data, an M&V analyst fits an energy prediction
model, using variables such as outdoor temperature and time
[@mathieu_quantifying_2011] to project baseline energy consumption in
the post-retrofit period. The difference between the counter-factual
baseline and the measured intervention period energy consumption
represents the energy savings. One of the key drawbacks of this method
is its reliance on a two-year timeline to quantify savings, during which
baseline measurements can become outdated due to changes in building
performance caused by non-routine events unrelated to the intervention.
Specifically, an M&V analyst should account for two types of baseline
changes: (1) static changes, which are known building operational
changes: such as renovation, equipment addition or removal, and changes
related to tenant turnover; and (2) gradual changes, which often occur
subtly and unknown, such as the incremental adoption of LED lighting or
more efficient plug loads. Identifying and quantifying these changes are
particularly labor-intensive and challenging. Consequently, those
limitations reduce the feasibility of M&V and complicate the
quantification of estimation uncertainty, thus impacting the accuracy
and timeliness of savings assessments.

@@

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston2019; @duan2013; @gupta2019;
@mie2022; @munro; @raseduzzaman2017]. An early application of this
randomized design to the field of M&V in buildings was first used in
[@raftery2018], which compared the energy performance of two different
supply air temperature reset strategies in a building. Subsequent
improvements to the M&V method include adding a sequential evaluation
framework and defining stopping criteria to end the M&V period early
after reaching a target level of uncertainty, using blocked
randomization by weekday. The full framework is detailed in a previously
published study [@raftery2024] along with a case study application to a
real building, and other example use cases. In summary, this method
provides M&V analysts with a randomized schedule that alternates between
baseline and intervention implementation while ensuring balanced
sampling across days of the week and seasons. For example, given a
10-week M&V period for one intervention, the balanced randomized
schedule would equally sample five Mondays with the baseline strategy in
operation and five Mondays with the intervention strategy in operation.
The major limitation of this method is that it is only applicable to a
subset of retrofit projects where interventions can be easily switched
on and off, which is a small subset of all the interventions performed
in buildings. However, for all applicable use cases, it allows analysts
to detect energy savings sequentially shortly after the test begins. In
addition, once the desired estimate uncertainty target is achieved,
analysts can terminate the M&V study and switch to 100% intervention.
The key advantage of randomization is that if control strategies are
sampled with equal probability, the influence of other confounding
factors such as occupancy change (a common non-routine event) and other
more subtle long term changes in building energy use are likely to be
evenly distributed among measurements, leading to a more accurate and
unbiased assessment of the intervention effect.

@@

Overall, those non-routine events mostly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. We found a similar concept briefly defined in
Guideline 14 as the 'load creep' [@ashrae_ashrae_2023] during multi-year
M&V project. However, currently there is no detailed instruction
provided. If M&V analysts are unaware of such changes and lack an
appropriate adjustment method (e.g., replacing filters before the
intervention begins), the savings could be overestimated as the
decreased energy use is incorrectly attributed to the intervention
rather than the filter replacement. An operational drift has not been
formally defined in the literature or standards, but in this study, we
consider it as a special type of non-routine event, as it also
represents a change in facility operation that occurs gradually over a
longer period of time.

@@

For all qualified buildings, we also create a feature to identify
buildings with very stable electricity usage overall by identifying
those with no statistically significant difference (P-value \> 0.05)
between their two-year electricity usage. Applying this filter to real
buildings, only 66 buildings are labeled in this 'stable' subset out of
a total of more than 600 buildings. While we did not adjust for weather
differences between the two years when generating this high-level
feature, this result highlights that the assumption that buildings are
typically stable over the timeframe involved in conventional M&V methods
is rarely true. As mentioned in \@ref(non-routine-events-impact), such
variability is largely associated with non-routine events, especially
operational drifts since the assessment, even under ideal situations,
lasts for two years. And according to the statistics shown here, it is a
typical case observed in real buildings applications.

@@

| Parameter              | Description                                                      | Value |
|-------------------|-----------------------------------|-------------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

-   A minimum and maximum of 12 and 108 weeks, respectively. The
    randomized schedule covers the entire two-year period but the
    stopping criteria enable an early stop at the end of the satisfied
    blocking period.

@@

The accuracy ($Acc$) is calculated as: $$
Acc = \frac{E_{baseline} - E_{intervention}}{E_{baseline}}-\frac{E'_{baseline} - E'_{intervention}}{E'_{baseline}}
$$where $E$ indicates true electricity usage ($E_{intervention}$ is the
'true' electricity usage calculated from Figure \@ref(fig:chwst) and
above equations) and $E'$ indicates the estimated electricity usage
either through the conventional method or the new randomized method. We
present the distribution of savings estimation accuracy using boxen
plots, also known as letter plots, which is an advanced variation of the
box plot designed to extend beyond the interquartile range (IQR) by
progressively dividing the data into smaller percentiles, revealing more
detail in the tails of the distribution. We set the division parameter
to k = 4, meaning the entire box area represents the data distribution
from 6.25% to 93.75%. As the steps move closer to the center line (50%
median), the distribution range progressively narrows with the next step
representing 12.5% to 87.5%, followed by 25% to 75%. In addition, for
each boxen plot, we also showed the mean value of the distribution on
top for reference.

@@

Additionally in the literature, we found another M&V study using over
500+ commercial buildings (on a different open-source dataset), but
their study scope was to compare the predictive performance of a variety
of M&V baseline models from simple weekly mean to more complex machine
learning methods using the normalized mean bias error (NMBE)
[@granderson_accuracy_2016]. Since we also calculated the normalized
savings estimation error, we compared our normalized accuracy for 600+
buildings with the TOWT model and TMY weather againts their TOWT model
prediction assessment in the following table, highlighting that the
model and overall building datasets are performing similarly.

|                 | Reference study results [@granderson_accuracy_2016] | Our study results      |
|-------------------|----------------------------------|-------------------|
| 25th percentile | -5.85                                               | `r percentiles$Q1`     |
| 50th percentile | -1.25                                               | `r percentiles$Median` |
| 75th percentile | 3.86                                                | `r percentiles$Q3`     |

@@

In some cases, particularly those involving large thermal mass and
active thermal energy storage systems or operating continuously, there
can be a substantial time-lagged effect associated with interventions.
This means that the effect of one control strategy may persist, or
'carryover', after switching to another. To mitigate such carryover
effects, building analysts may choose to drop non-consecutive days in
the switchback experiment. For example, consider a heat pump water
heater where the tank is charged at very different times of the day by
the intervention strategy compared to the baseline strategy. When
switching strategy, the tank may be much warmer or cooler than typical
for that strategy, thus influencing subsequent measurements,
particularly the performance measured on the first day (i.e., a
non-consecutive day). In such cases, increasing the sampling interval
(e.g., sample strategies every 2 or 3 days instead of daily) and
discarding measurements from those non-consecutive days can eliminate
carryover. However, this approach reduces the total data collected: for
instance, sampling every 3 days and dropping days with non-consecutive
control strategies would drop 1/6 of measurements on average. Sampling
every 2 days means dropping a quarter of the measurements. Therefore, by
increasing the sampling interval, there is an accuracy penalty due to a
reduction in randomization (how often samples occur), and when
decreasing the sampling interval, there is an accuracy penalty from
dropping more non-consecutive days. The dataset contains no specific
information about the buildings' thermal mass condition, but to quantify
such trade-off, we compared the accuracy of different sampling intervals
and the consequence of dropping the non-consecutive days at the time
when randomized M&V produces a result (after satisfying all stopping
criteria) in Figure \@ref(fig:absinterval).

@@

```{r abscont, fig.cap="Comparison of different sampling ratio impact on M&V estimation accuracy over the entire 24 months (previous results of sampling at 50%/50% shown on the left side of the red dashed line)", fig.height=8, fig.width=12}

@@

In this study, we set the baseline period to begin on January 1st, which
offers the advantage of capturing colder outdoor weather conditions
early in the M&V process. However, when the baseline measurements miss
the initial winter period, meeting the required weather condition
criteria can potentially take a longer time. A previous study also
showed that M&V accuracy, especially model fitting, depends on both
monitoring duration and starting month, with optimal modeling achieved
when sufficient weather variation is captured [@singh_predicting_2014].
To assess this impact in our switchback randomization, we repeated the
randomized M&V analysis for all buildings using a new start date in
March. As shown in Figure \@ref(fig:startT), the new M&V process results
in a noticeable delay in satisfying the stopping criteria. Compared to
Figure \@ref(fig:intervaltime), where a January start allowed most
buildings to finish within 36 weeks, a March start extended the required
duration to 48 weeks for the majority of buildings. Furthermore, when a
longer sampling interval is used or non-consecutive days are excluded
(to avoid carryover effects), some buildings may require up to 60 weeks
to complete the randomized M&V. As expected, to collect sufficient data
representing winter conditions, the sampling may need to continue until
the next January (around week 48). However, as the overall data
collection period extends, the influence of dropping non-consecutive
days diminishes and becomes less of a concern.

@@

Specifically, we showed that throughout 11 different locations assessed
in this study, the randomized M&V can provide a savings estimation after
36 weeks (with the majority finished by 24 weeks) once all stopping
criteria are satisfied. In contrast, the conventional method requires a
full range of baseline and intervention measurements under normal
operating conditions, typically taking 6 - 9 months for each phase. Most
importantly, we verified that with a reduced M&V timeline, the
randomized method can estimate savings more accurately by showing that
the absolute error is only 1 - 2% for a typical building. Whereas for
the much longer two-year conventional method, the estimation error was
approximately 5% in a typical building. We also evaluated the impact of
non-routine events, especially more gradual and subtle operational drift
on the proposed M&V method and the results show that baseline changes in
the post-retrofit period can deviate savings estimated using the
conventional method. In contrast, we found that those events have a very
negligible impact on the savings estimated using the randomized method,
demonstrating robustness to this issue. Furthermore, we consider
scenarios where there is a known carryover effect from switching between
strategies in the building and assess the impact of dropping samples
when the strategy is non-consecutive and varying the sampling interval.
Through that process, we showed that for the typical building, if
carryover is present, using a 2-day sampling interval and dropping days
with non-consecutive control strategies operating yields the optimal
design for the median building in this dataset. Lastly, we investigated
the time required to complete the randomized M&V method under different
scenarios. We found that measuring a sufficient range of outdoor weather
conditions remains the most stringent criterion. However, even in the
most non-ideal case, the randomized method can still produce an estimate
much faster than two years, and produces an estimate with far lower
error.
